### 2021-11-10 NSF2021/Description/specificaim1.tex
RJR - SP1, Prelim, new figure folders

\section{Proposed Research}
\label{sec:proposed-work}
%%
We have demonstrated that our hydrophobic attraction with repulsion
potential (HAP) approach efficiently simulates self-assembly of
amphiphilic particles into two-dimensional micelles, bilayer membranes,
and vesicles \cite{Fu2018_SIAM} and recreates the tank-treading
phenomenon in external shear flows \cite{FuQuRyYo20}.
While the results show great promise in the field of collective body
hydrodynamics, several outstanding issues need to be addressed. These
include a thorough analysis of elastic properties of our coarse-grained
bilayers and efficiently simulating three-dimensional collective
hydrodynamics of amphiphilic particles.
We must also incorporate electric charge and account
for how external fields control particle self-assembly.

@@

The goal of Specific Aim 1 is to investigate generalizations of the
screened Laplace equation for different soft matter systems.

\begin{itemize}
  \item Include ions in suspension
  \item Discussed in Sylvie Roke's talk---I added the slides to the repo
\end{itemize}

\begin{comment}
The goal of Specific Aim 1 is to characterize the material properties of
many-body, self-assembled amphiphiles.  For amphiphiles assembled into
bilayers, these properties are described by membrane continuum
mechanics.  Our goal is to map the parameters of the particle-based
model onto the elastic moduli from continuum theory.  Results from this
goal will facilitate simulators to use the hydrophobic attraction force
calculations to model bilayers with specific composition. These
calculations have provably less computational complexity than those of
molecular dynamics simulations and possess the molecular granularity
lacking from continuum models.

>>>>

\section{Proposed Research}
\label{sec:proposed-work}
%%
\begin{center}
  \begin{figure}
      \includegraphics[width=1\textwidth]{figures/SpecificAim1/Domain.pdf}
  \end{figure}
\end{center}

@@

The proposed research aims to provide fundamental insight into the
self-assembly dynamics of amphiphilic particles. These results will
facilitate optimal design of smart materials by tuning the geometry and
properties of the amphiphilic particles.
Specific Aim 1 achieves this through a novel mathematical model
for particle self-organization driven by long-range interfacial forces.

\subsubsection{Problem formulation}
The mathematical formulation is a nonlinear system 
for the dynamics of a collection of particles. The 
interactions come from a system of linear partial differential equations
(PDE) and comprise hydrodynamic interactions and hydrophobic interactions.
The formulation prevents particle collisions through excluded volume
potentials.

The hydrodynamic interactions come from the mobility problem for rigid
particles immersed in a viscous solvent. Assuming inertial terms are
negligible, the particle motion is goverened by the Stokes equations
\begin{alignat}{3}
\label{eq:stokes1}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, 
    && \xx \in \Omega, \\
\label{eq:stokes2}
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
\label{eq:stokes3}
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity and $p$ is the pressure of the solvent,
$\uu_\infty$ is the background flow, and $\mu$ is the constant
solvent viscosity. The domain
$\Omega$ is the fluid region surrounding the particles.
Since the particles are rigid, the solvent velocity 
satisfies a no-slip boundary condition for a rigid body motion 
%\begin{align}
%\label{eq:rigid_bc}
%  \uu(\xx) = \vv_i + \omega_i  (\xx - \aa_i)^\perp, \quad
%    \xx \in \Gamma_i,\qquad  i=1,\ldots,N_p,
%\end{align}
on the particle boundary $\Gamma_i$
with translational velocity $\vv_i$ and
angular velocity $\omega_i$.
Given imposed forces acting on each particle,
the \emph{mobility problem} consists of finding
translational and angular velocities so that 
viscous fluid forces balance the imposed forces.

The hydrophobic interactions come from the tendency 
of particles to minimize the free energy of the structure of
the surrounding water molecules.  
Experimental measurements suggest that the free energy functional
take the form 
\begin{equation}
\label{eq:free_energy}
F[u] = C \int_{\Omega}  \rho |\nabla u|^2 + \rho^{-1} f(u) \,dx
\end{equation}
where
$u$ is an order parameter for the structure of water,
$\rho$ is a decay length,
$C$ is a constant, and $f(u)$ is a potential.
Hydrogen-bond persistence times are on the order of picoseconds \cite{MaGa13}.
Since waters reorient 
on a time scale much smaller than for the particle motion, 
then at each moment in time $u$ minimizes $F[u]$ and satisfies 
\begin{equation}
\label{eq:SL}
-\rho^2 \Delta u + f'(u) = 0  \text{ in } \Omega,\quad u = g
\text{ on } \Omega.
\end{equation}
The boundary condition $g$ that imposes 
the water structure at the water-hydrocarbon interfaces.
It is a material label that is transported with the particle motions.

The coupling between the Stokes equations \eqref{eq:stokes1}-\eqref{eq:stokes3}
and the semilinear elliptic equation \eqref{eq:SL}
comes from the first variation of the energy functional.
The particles minimize the free energy $F[u]$
by altering the location of hydrophobic interfaces. 
It is possible to calculate the rate of change of $F[u]$
using variation of the domain~\cite{Bandle2015, Schiffer1954, Grinfeld2010}.
Carrying out this variation yields a 
%The hydrophobic forces are torques are
%\begin{equation}
%  \label{eq:hydrophobicAttraction}
%  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu \, \dif s, 
%    \quad 
%    T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot
%    ({\bf T} \cdot \nnu) \dif s.
%\end{equation}
surface force density 
\begin{align}
  \label{eq:stress}
\mathbf{T}
= C \left[ \rho^{-1} f(u) \mathbf{I}
  + \rho  \left(|\nabla u|^2 \mathbf{I} - \nabla u  \nabla u^T\right)\right].
\end{align}
The imposed forces and torques from the integration of $\mathbf{T}$ along the
particle boundary.
The density \eqref{eq:stress}  was first introduced
by~\cite{Fu2018_SIAM} and is the higher-dimensional analogue of
disjoining pressure calculated by
\cite{MaRa76, ErLjCl89, KoNa15, Nagle17, KUZMIN2005}.
It is the mathematical ingredient responsible for forming particle
aggregates that sequester their hydrophobic surface regions.

By solving the above system, we obtain translational and angular
velocities of the many-body system. 
A second-order Adams-Bashforth scheme updates the particle positions and orientations.

The formulation presents a number of
mathematical and numerical challenges. For one, the domain is constantly
changing so that a new boundary value problem must be solved at each time step.
The particle boundaries are nearly touching and this requires
high-order numerical schemes to accurately calculated forces.
The domain is unbounded and there are issues with accurately
calculating the velocity in the far-field. 
Finally, the numerical routine must handle
large system sizes efficiently when their are many particles
is large.

%and
%\begin{equation}
%  \label{eq:force}
%  \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \FF_i,\quad 
%  \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot (\ssigma \cdot \nnu) \,
%  \dif s = T_i, \qquad i=1,\ldots,N_b,
%\end{equation}
%where
%$\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T \right)$
%is the hydrodynamic stress tensor (pressure tensor) and $\nnu$ is the
%particle outward normal.

\subsubsection{Janus-particle self-assembly} 
\begin{wrapfigure}[14]{l}{0.5\textwidth}
\includegraphics[width=0.5\textwidth]{figures/SpecificAim1/multilamelar.pdf}
\caption{\label{fig:multilamelar}
  Experimentalists can transform unilamellar vesicles to multilamellar
  vesicles by altering the quality of the solvent
    \cite{deMeetal21}.
  In \eqref{eq:SL}, the solvent quality enters through the potential
  $f(u)$ and boundary conditions. 
}
\end{wrapfigure}
A number of morphologies arise from the theory. In the linear response case,
take $f(u) = u^2$ so that water structure has its lowest potential value
at $u = 0$, representing bulk water. The solutions of \eqref{eq:SL} have
a boundary layer structure and decay to 0 in the bulk with the decay
length $\rho$.

Consider a polar boundary condition where
$g_i(\mathbf{x}) = (\pi r)^{-1/2} \cos (\theta_i(\mathbf{x}))$
where $\theta_i$ is the angle formed by $\mathbf{x}$, the particle
center $\mathbf{a}_i$, and the particle director $\mathbf{d}_i$.
The normalization $(\pi r)^{-1/2}$ here and others below
provide for a uniform surface energy $\int_{\Gamma_i} g_i^2 \,ds = 1$
when the particles are disks of radius $r$.
The boundary condition is odd along the particle axis 
making the particle head repel the tail.
We simulate the dynamics of 100 particles with random initial position
and orientation. The particles rapidly form chains with their
directors perpendicular to the length of the chain. The equilibrium
structure consists particles positioned on a square grid with alternating
orientations.
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A5.png}\\
    Figure 2. Self assembly for a polar boundary condition
    \label{fig:self-assemblyA}
\end{center}

Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and are accounted for as follows.
The hydrophilic side takes the value $u = 0$ since the
apolar headgroup of a lipid, for example, does not alter the structure
of adjacent waters. The hydrophobic side takes the value $u > 0$. 
The interaction between particles is attractive. Using
$(3\pi )^{-1/2}(1 + \cos(\theta_i(\mathbf{x})))$
as a simple modification to the previous boundary condition 
produces two-dimensional micelles and bilayers \cite{Fu2018_SIAM}.
Researchers have asked us if the phenomenon is robust. To show
that it is, we use 
\begin{equation}
\label{eq:vonMises}
g_i(\mathbf{x}) =
\sqrt{
  \frac{\exp( \kappa_i(\theta_i(\mathbf{x}) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}
\end{equation}
instead where the radicand is the von Mises 
distribution with director angle $\theta^0_i$ and Bessel function
$I_0$ of order zero. The square of this function behaves as a normal
distribution with variance like $1/\kappa_i$. To replicate
experimental conditions, the concentration $\kappa_i$ and radii
$r_i$ are randomized. Using the same initial configuration, the
particles form small groups. The small groups combine to form bilayers.
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B4.png}\\
    Figure 3. Self assembly for a amphiphilic boundary condition
    \label{fig:self-assemblyA}
\end{center}

Multilamelar bilayers arise with an increase by shifting the boundary condition
further. Let $g_i(\mathbf{x}) = (9\pi )^{-1/2}(2 + \cos(\theta_i(\mathbf{x})))$.
This gives a hydrophobic particle with greater intensity of hydrophobicity
at $\theta_i = 0$ than at $\theta_i = \pi$.  The initial self-assembly
is similar to that of the amphiphilic particle case, except that the bilayers
do not stay well separated.  Rather, the bilayers stack on top of each
other as a consequence of the interfacial tension of exposed particle heads. 
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C4.png}\\
    Figure 3. Self assembly for hydrophobic particles with bias.
    \label{fig:self-assemblyA}
\end{center}

In summary, the hydrophibic interaction-mobility problem model yields
rich morphologies that are valuable in the synthesizing multiphasic and
anisotropic colloids. The goal is to obtain numerically aided 
syntheses and assembly of soft matter materials.
Applications in this area include cargo-carrying capacity, 
catalytic activity, or stimuli-responsive shape-change to amplify colloid
utility for a wide range of  spanning environmental remediation to drug
delivery. The main experimental challenges include the ability to
decouple particle surface properties.
The main modeling challenge is to translate the phenomenological
parameters, e.g. boundary conditions, into measurable quantities
such as contact angles. Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions. In Specific Aim 3, we
discuss how to include ions and surface charge.

\subsubsection{Hydrophobic forces as capillary wetting}
A natural generalization of the linear response model is 
to use a fourth-order, tilted double-well potential
$f(u) = a(u-u_0)^2(u-u_1)^2 + bu$.
Motivated by capillary wetting, soft matter physicist Gerhard Gompper 
introduced a double-well model for one-dimensional hydrophobic attraction
\cite{GoHaKo94}. The two-dimensional attraction we propose
is more complicated and includes extra line tension effects
not present in the one-dimensional theories.

Here, the order parameter $u$ is the mean number of nearest
neighbors per water molecule.
The value $u_0 = 4$ gives a local minimum for the ideal
tetrahadral water network found at hydrophobic surfaces.  
The value $u_1 \sim 4.4$ gives the absolute minimum
in the bulk where there is a mixture of four and five-coordinated
waters. The equation \eqref{eq:SL} takes the form of a
steady-state Allen-Cahn equation. 
When particles are well-separated, the energetic cost for filling
the interstitial region with $u_0$-phase water is prohibitive
and there is a phase shift from the $u_0$ to $u_1$ water order
---the attraction landscape is more
or less identical to the linear response theory. Conversely, when
particles are proximal, it becomes energetically favorable
to coalesce $u_0$-phases. This introduces a line tension
for the phase transition.

One potential application of the Allen-Cahn-type formulation is to
explaning measured hydrophobic attraction profiles.
The experimental data on hydrophobic attraction between 
is well-established and the literature contains data for a number of
solvents and hydrocarbon species. The experimental measurements
are well-fit by a double exponential curve in the 10 to 100 nanometer range
of separation.  Below 10 nm, however, there is an amplification of force
and the attraction follows a much steeper force law not accounted
for by past phenomenoloical theories \cite{Lin2005}.

It may be the case that the double-well formulation recreates the
long and short-range interactions.  The advantage
is that the two-dimensional simulations can replicate the
experimental setup better than prior, one-dimensional theoretical work 
Some of the model parameters directly connect 
to long-range decay lengths and energy cost of creating hydrophobic-water
interfaces with established ranges of values. Other parameters, like the
well structure, are fit to the measured data. These base-line simulations
will allow us to move on to provide more accurate colloidal suspension
simulations.

Other approaches will be required if the modified phenomenological theory 
does not recreate the measured force curves.
We can rule out convection in the water structure since  
the time scale of particle motion is much larger than
the hydrogen-bond lifetimes. One approach is to consider Gaussian fluctuation.
Here, we expand the free energy functional around the saddle point solution.
The fluctuation energy then involves calculation of the eigenfunctions of
the second variation of the free energy. This accounts for the fact
that experimental measurements are carried out at finite 
temperatures, where thermal fluctuations should ideally also be considered.

%
% The following commented out Nov 6, 2021.
%
%
\begin{comment}
The goal of Specific Aim 1 is to characterize the material properties of
many-body, self-assembled amphiphiles.  For amphiphiles assembled into
bilayers, these properties are described by membrane continuum
mechanics.  Our goal is to map the parameters of the particle-based
model onto the elastic moduli from continuum theory.  Results from this
goal will facilitate simulators to use the hydrophobic attraction force
calculations to model bilayers with specific composition. These
calculations have provably less computational complexity than those of
molecular dynamics simulations and possess the molecular granularity
lacking from continuum models.
